---
title: "Your Title"
author: "Your Name"
format: 
  html:
    embed-resources: true
---


```{r}

```

```{r}
library(tidyverse)
stat113_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/stat113.csv")
stat113_df

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_histogram()

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_histogram(colour = "black", fill = "white", bins = 17)
```

## Section 2.4: Practice Exercises

```{r}
stat113_restructured <- stat113_df |> group_by(SocialMedia) |>
  summarise(n_social = n())
stat113_restructured

## b
ggplot(data = stat113_restructured,
       aes(x = SocialMedia)) +
  geom_bar()

## c
ggplot(data = stat113_restructured,
       aes(x = SocialMedia, y = n_social)) +
  geom_col()
```

Class Exercisse 2

```{r}

fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 TRUE ~ "weekday"))
ggplot(data = fitness_full, aes(x = active_cals, y = distance)) +
  geom_point() +
  geom_smooth(se = FALSE, span = 0.00001)
## method = "lm"
## span controls how "wiggly" the smoother is
```

Class Exercise 3

```{r}
fitness_full
ggplot(data = fitness_full, aes(x = Start, y = steps)) +
  geom_line() +
  geom_smooth() 
```


facet_grid(name_of_var ~ name_of_var2)

```{r}
ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_grid(Sport ~ Sex)

ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_wrap(Sport ~ Sex)
```

Class Exercise 5

```{r}
ggplot(data = stat113_df, aes(x = Sport, y = GPA)) +
  geom_boxplot(outlier.shape = 8)
```

Class Exercise 6

```{r}
ggplot(data = stat113_df, aes(x = Haircut,
                              y = ..density..)) +
  geom_histogram(colour = "black", fill = "white")
```

Your Turn 1

done by hand

Your Turn 2

```{r}
ggplot(data = stat113_df, aes(x = GPA, y = Exercise)) +
  geom_point(aes(colour = Sport))

ggplot(data = stat113_df, aes(x = GPA, y = Exercise)) +
  geom_point() +
  facet_wrap(~ Sport)
```

Your Turn 4

```{r}
fitness_full
ggplot(data = fitness_full, aes(x = steps, colour = weekday)) +
  geom_freqpoly()

ggplot(data = fitness_full, aes(x = steps)) +
  geom_histogram() +
  facet_wrap(~ weekday)
```


Your Turn 5

```{r}
ggplot(data= stat113_df, aes(x = GPA, y = Exercise)) +
  geom_point()
```

Line plots only make sense when there is one y value per unique x value.

Your Turn 6

There are 70 NA values of GPA, likely from first year students filling out the survey.

Your Turn 7 and 8

Skip

## Section 2.5

## Section 2.6

Class Exercise 1.

```{r}
ggplot(data = stat113_df, aes(x = Wgt, y = Hgt,
                              colour = Sport)) +
  geom_point(colour = "purple") +
  geom_smooth()
```

```{r}
ggplot(data = stat113_df, aes(x = TV)) +
  geom_histogram(colour = "black", fill = "white")
```

Your Turn 1

Both b and e would work. These are the only two sets of code where purple is not in an `aes()` function and where `groupvar` is in an `aes()` function.

Your Turn 2

If specified locally, `aes(x = xvar, y = yvar)` would have to go in both `geom_point()` and in `geom_smooth()`. Because usually they are used in the remaining geoms, `x` and `y` aesthetics are usually specified globally.

```{r}
stat_social <- stat113_df |> group_by(SocialMedia) |>
  summarise(n_social = n()) |>
  mutate(SocialMedia = fct_reorder(SocialMedia, n_social))

ggplot(data = stat_social, aes(x = SocialMedia,
                               y = n_social)) +
  geom_point() +
  geom_segment(aes(xend = SocialMedia, y = 0,
                   yend = n_social)) +
  coord_flip()
```
