---
title: "Section 2: ggplot"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---



```{r}
library(tidyverse)
stat113_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/stat113.csv")

stat113_df
names(stat113_df)

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_histogram(colour = "black", fill = "white",
                 bins = 17)

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_freqpoly()

ggplot(data = stat113_df, aes(x = GPA)) +
    geom_histogram(colour = "black", fill = "white", bins = 17) +
    geom_freqpoly(bins = 15)
```

Exercise 1.

```{r}
## names(stat113_df) shows the names of the data frame

## exercise 2
ggplot(data = stat113_df, aes(x = GPA)) +
    geom_histogram(colour = "dodgerblue4",
                   fill = "dodgerblue1", bins = 17)

## exercise 3
ggplot(data = stat113_df, aes(x = GPA)) +
    geom_histogram(colour = "chartreuse",
                   fill = "yellow4",
                   bins = 17)
```

```{r}
ggplot(data = stat113_df, aes(x = SocialMedia)) +
  geom_bar()
```

Exercise 4, Exercise n5.

```{r}
ggplot(data = stat113_df, aes(x = Award)) +
  geom_bar(colour = "chartreuse",
           fill = "yellow4")
```

Exercise 6.

```{r}
ggplot(data = stat113_df, aes(x = Sex, fill = Sport)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion")
## in this data set, a slightly higher proportion
## of males play a sport than females do.
```

```{r}
ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_boxplot()
```

Exercise 7 and 8.

```{r}
ggplot(data = stat113_df, aes(x = Year, y = GPA)) +
  geom_violin(colour = "lightpink4", fill = "lightpink")
```

## Class Exercises

Class Exercise 1.

```{r}
stat113_df

ggplot(data = stat113_df, aes(x = SocialMedia)) +
  geom_bar()

stat113_restructured <- stat113_df |> group_by(SocialMedia) |>
  summarise(n_social = n())
stat113_restructured

ggplot(data = stat113_restructured, aes(x = SocialMedia)) +
  geom_bar()

ggplot(data = stat113_restructured,
       aes(x = SocialMedia, y = n_social)) +
  geom_col()
```

Class Exercise 2.

```{r}
library(tidyverse)
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 TRUE ~ "weekday"))
```

```{r}
ggplot(data = fitness_full, aes(x = distance, y = active_cals)) +
  geom_point() +
  geom_smooth(span = 0.1, se = FALSE,
              method = "lm")
## span controls how wiggly the smoother is
## method = "lm" fits a linear regression model.
```

Class Exercise 3. 

```{r}
fitness_full
ggplot(data = fitness_full, aes(x = Start,
                                y = steps)) +
  geom_line() +
  geom_smooth()
```

Class Exercises 4 and 5.

```{r}
ggplot(data = stat113_df, aes(x = Exercise,
                              y = GPA)) +
  geom_point() +
  facet_grid(Sex ~ Sport)

ggplot(data = stat113_df, aes(x = Exercise,
                              y = GPA)) +
  geom_point() +
  facet_wrap(~ Sex + Sport) ## not as nice

ggplot(data = stat113_df, aes(x = Award,
                              y = Exercise)) +
  geom_boxplot(outlier.shape = 8)
```

Class Exercise 6. Skip.

## Your Turn 

Your Turn 1. See Handout

Your Turn 2 and 3.

```{r}
ggplot(data = stat113_df, 
       aes(x = Haircut, y = GPA)) +
  geom_point(aes(colour = Sport))

ggplot(data = stat113_df, 
       aes(x = Haircut, y = GPA)) +
  geom_point() +
  facet_wrap(~ Sport)
```

Your Turn 4

```{r}
ggplot(data = fitness_full,
       aes(x = steps)) +
  geom_freqpoly(aes(colour = weekday))
## or, equivalenttly, with a global 
## colour aesthetic:
ggplot(data = fitness_full,
       aes(x = steps, colour = weekday)) +
  geom_freqpoly()

ggplot(data = fitness_full,
       aes(x = steps)) +
  geom_histogram() +
  facet_wrap(~ weekday)
```

Your Turn 5

A line plot is only useful if there is one value of y for each uniquye value of x.

```{r}
ggplot(data = stat113_df,
       aes(x = GPA, y = Exercise)) +
  geom_line()
## this doesn't make sense to make
## because there can be many students
## with one GPA
```

Your Turn 6

Those 70 values of GPA are missing, likely from first-years in STAT 113. 

## Section 2.5



## Section 2.6

Class Exercises

Class Exercise 1.

```{r}
ggplot(data = stat113_df,
       aes(x = Wgt, y = Hgt)) +
  geom_point(aes(colour = Sport)) +
  geom_smooth()
```

Points are now coloured but the smoother does not use sport as a colour aesthetic so there is just one smoother now.

```{r}
## class exercise 2
ggplot(data = stat113_df, aes(x = TV)) +
  geom_histogram(colour = "black", fill = "white")
```

Your Turn 1.

a will not work because "lm" is not a variable in our data set (so should not go inside aes()).


```{r}
ggplot(data = stat113_df,
       aes(x = Wgt, y = Hgt, colour = Sport)) +
  geom_point(colour = "purple") +
  geom_smooth()
```

b will work: specifying `colour = "purple"` as a local option to geom_point() overrides the global aesthetic.

c will not work: groupvar should be an aesthetic while purple should not be.

d will not work: groupvar should be an aesthetic.

e works!

f does not work: purple is not a variable so should not go inside aes().