---
title: "Section 3: Wrangling with `dplyr`"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

## Section 3.2

Class Exercises

Class Exercises 1 and 2.

```{r}
#| echo: false
library(babynames)
library(tidyverse)

babynames_top10m <- babynames |>
  filter(year == 2017 & sex == "M") |>
  arrange(desc(n)) |>
  slice(1:10)

ggplot(data = babynames_top10m, aes(x = name,
                                    y = n)) +
  geom_col()
```

Class Exercises 3 and 4

```{r}
babynames_jack <- babynames |>
  filter(name == "Jack" & sex == "M")

ggplot(data = babynames_jack, aes(x = year,
                                  y = n)) +
  geom_line()
```

Class Exercise 5.

When piping with data frames, the first argument of the function we use must be for a data frame (not a vector like it is for `mean()` and not a formula like it is for `lm()`).

Your Turns

Your Turn 1 and 2.

```{r}
## a
babynames |> filter(year != 1945)

## b
babynames |> filter(n >= 400 & n <= 5000)

## c
babynames_hm <- babynames |>
  filter((name == "Monica" & sex == "F") |
                      (name == "Hilary" & sex == "F"))

ggplot(babynames_hm, aes(x = year, y = n)) +
  geom_line(aes(colour = name))
```

Your Turns 3 and 4.

```{r}
babynames_5 <- babynames |> 
  filter((name == "Matthew" & sex == "M") |
           (name == "Ivan" & sex == "M") |
           (name == "Jessica" & sex == "F") |
           (name == "Lillian" & sex == "F") |
           (name == "Robin" & sex == "M"))
babynames_5
ggplot(data = babynames_5, aes(x = year,
                               y = n)) +
  geom_line(aes(colour = name))

## Your Turn 4

babynames_2002 <- babynames |>
  filter(year == 2002 & sex == "F") |>
  arrange(desc(prop)) |>
  slice(1:10)
ggplot(data = babynames_2002, aes(x = name, y = prop)) +
  geom_col(colour = "darkred", fill = "white")

```





## Section 3.4: Practice

Class Exercises 

Class Exercise 1. 50% might make more sense if you are going by a strict "majority" definition. But 55% makes more sense if you're thinking about this more as "are females represented" in each of the majors?



Class Exercises 4 and 5

```{r}
babynames |> group_by(year, sex) |>
  mutate(rank_var = rank(desc(n))) |>
  filter(rank_var == 1) |>
  ungroup() |>
  group_by(name) |>
  summarise(n_appear = n()) |>
  arrange(desc(n_appear))
```

Class Exercise 6

```{r}
babynames_test <- babynames |>
  group_by(year, sex) |> mutate(ntest = n / prop)
babynames_test |> slice(1, 2, 3, 4, 5)
## since we have grouped by year and sex,
## slice(1, 2, 3, 4, 5) is giving the first 5 rows
## __of each group__.
babynames_test |> ungroup() |> slice(1, 2, 3, 4, 5)
```

Your Turn

Your Turns 1 through 3
```{r}
toy_df <- tibble(x = c(NA, 3, 4, 7),
                 y = c(1, 4, 3, 2),
                 z = c("A", "A", "B", NA))
toy_df

## Your Turn 1
## a: mutate
toy_df |> group_by(z) |>
  mutate(rank_y = mean(y))
## calculating mean fopr each level of z

## b: slice
toy_df |> group_by(z) |>
  slice(1)

## c: filter
toy_df |> group_by(z) |>
  filter(x > 3.5) ## the same output as if we did not group_by() z.
toy_df |> group_by(z) |>
  filter(y == min(y)) ## does matter if we use a summary function inside filter()

## d: arrange
toy_df |> group_by(z) |>
  arrange(x)
## arrange() is not affected by group_by()

## e: select
toy_df |> group_by(z) |>
  select(2)
## select() on a grouped data frame always
## returns the group variable.

## Your Turn 2
## a: mutate
toy_df |> mutate(x_over_y = x / y)
## x_over_y is NA for any row with missing x or y

## b: slice
toy_df |> slice(1:2)
## NA's do not affect behavior of slice()

## b: arrange
toy_df |> arrange(x)
toy_df |> arrange(desc(x))
## NA gets put at the end (no matter what order you sort by).

## c: filter
toy_df |> filter(x > 5)
toy_df |> filter(x <= 5)
## NA values are removed for these types of filter()s.

## d: summarise
toy_df |> summarise(mean_x = mean(x))
## returns a missing value if ANY value of x is NA


## e
toy_df |> group_by(z) |>
  summarise(mean_x = mean(x))
## mean is missing for any level of the grouping
## variable z that has a missing x value
```

Your Turn 3

a. mutate is generally safe: just adding a variable but the rest of the data set would still be "okay" even if you made a mistake.

b. arrange is generally safe: you're just sorting the data set.

c. filter is generally not as safe: you would have to be sure that you never need access to the rows you filter out again.

d. summarise is generally not as safe: you're changing the structure of the data to be only one row per group (and you lose access to the original data).

e. select is generally not as safe: you would have to be sure that you never need access to the columns you select out again.

Your Turn 4

```{r}
slumajors_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/SLU_Majors_17_21.csv")
slumajors_df <- slumajors_df |>
  mutate(ntotal = nfemales + nmales) |>
  mutate(percfemale = 100 * nfemales / (nfemales + nmales))
slumajors_df |>
  summarise(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) 
## summarise() collapses the data set so that there
## is one row per group (here there are no groups so
## there is just one row). 
slumajors_df |>
  mutate(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) |>
  relocate(meantotalmajor)
## relocate(meantotalmajor) is equivalent to
## select(meantotalmajor, everything())

## mutate() just adds a new variable (keeps the same number of rows and all of the other variables)
```
Your Turns 5 through 7


```{r}
##babynames |> group_by(year) |>
##  mutate(n_unique = n())
babynames_unique <- babynames |> group_by(year) |>
  summarise(n_unique = n())
## summarise more convenient here than mutate because
## it lets us get one row per year
ggplot(babynames_unique, aes(x = year, y = n_unique)) +
  geom_line()
## since 1880, the number of unqiue names in the US has 
## increased (with a little drop at the end)

## Your Turn 6
babynames_unique_ys <- babynames |> group_by(year, sex) |>
  summarise(n_unique = n())

ggplot(babynames_unique_ys, aes(x = year, y = n_unique)) +
  geom_line(aes(colour = sex))
```

Your Turn 7

```{r}
babynames |> group_by(name) |>
  summarise(n_births = sum(n)) |>
  arrange(desc(n_births))
```
