---
title: "Section 4: Communication"
author: "Matt Higham"
execute: 
  output: false
  echo: false
  warning: false
format: 
  html:
    embed-resources: true
---

## Class Exercises

Class Exercise 1 - 6 (skip class exercise 5).

```{r}
library(gapminder)
library(tidyverse)
gapminder

country_df <- gapminder
country_df |> filter(continent == "Africa") |>
  filter(year == 1992 & lifeExp < 35)
country_df |> filter(continent == "Asia") |>
  filter(year == 1977 & lifeExp < 35)

label_df <- country_df |> filter(country == "Rwanda" |
                       country == "Cambodia") |>
  group_by(country) |>
  filter(lifeExp == min(lifeExp))
label_df
```

```{r}
#| output: true
#| fig-height: 4
ggplot(data = country_df, aes(x = year,
                              y = lifeExp,
                              colour = continent)) +
  geom_line(aes(group = country), alpha = 0.4) +
  facet_wrap(~ continent) +
  geom_text(data = label_df, aes(label = country),
            show.legend = FALSE) +
  scale_colour_viridis_d() +
  labs(x = "Year",
       y = "Life Expectancy",
       colour = "Continent",
       title = "Life Expectancy through Time",
       caption = "Rwandan Mass Genocide, Cambodian Civil War")

```

Skip Class Exercise 7.

Class Exercise 8.

## Your Turn

Your Turn 1.

The boxplot options have been changed with point-and-click so trying to recreate them exactly would be really challenging.

The removal of outliers is not reproducible unless he explicitly states in a write-up which outliers were removed.
Your Turn 2.

```{r}
library(tidyverse)
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend", 
                                 TRUE ~ "weekday"))
```

```{r}
fitness_label <- fitness_full |> filter(Start == "2019-04-14" |
                         Start == "2021-05-24") |>
  mutate(event = case_when(month == "Apr" ~ "half-marathon",
                           month == "May" ~ "grand canyon")) |>
  relocate(event) |> ## puts event as the first variable
  mutate(event_take2 = c("half-marathon", "grand canyon"))
ggplot(data = fitness_full, aes(x = Start,
                                y = steps)) +
  geom_line() +
  geom_smooth() +
  geom_text(data = fitness_label, aes(label = event))
```

Your Turn 3

```{r}
mtcars <- mtcars |> rownames_to_column() |> rename(car_name = rowname)
mtcars_label <- mtcars |> filter(mpg == max(mpg))

ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_text(data = mtcars_label, aes(label = car_name))
```

Your Turn 4.

```{r}
library(tidyverse)
hpi_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/hpi-tidy.csv")

ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears,
                          colour = Region)) +
  geom_point() +
  theme(panel.grid.major = element_line(colour = "green",
                                        size = 5),
        plot.background = element_rect(fill = "yellow4"),
        axis.text = element_text(colour = "yellow"),
        axis.line = element_line(linewidth = 6,
                                 colour = "salmon3"),
        legend.background = element_rect(fill = "green3"))
## ggplot plots are made up of rectangles, lines, and text,
## which can be modified in theme() with element_rect(),
## element_line(), and element_text()

## element_blank() can be used to get rid of something
## entirely
```