---
title: "Section 8: Base R"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
fitness_df <- read_csv(here("data/higham_fitness_notclean.csv"))
```

```{r}
weekend_vec <- c("Sat", "Sun")
weekday_vec <- c("Mon", "Tue", "Wed", "Thu", "Fri")

fitness_df |>
  mutate(weekend_ind = case_when(weekday %in% weekend_vec ~ "weekend",
                                 weekday %in% weekday_vec ~ "weekday")) |>
  relocate(weekend_ind)

## equivalent but needs more syntax
fitness_df |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 weekday == "Mon" | weekday == "Tue" | weekday == "Wed" | weekday == "Thu" | weekday == "Fri" ~ "weekday")) |>
  relocate(weekend_ind)

## var %in% c("A", "B", "C", "E") <==> var == "A" | var == "B" | var == "C" | var == "E"
```

Class Exercise 3.

```{r}
grades_df <- structure(list(section = c(1030, 1030, 1030, 1030, 1030, 1030, 
1030, 1030, 1030, 1030, 1030, 1030, 1030, 1030, 1030, 1030, 1030, 
1030, 1030, 1030, 1030, 1030, 1030, 1030, 1030, 1140, 1140, 1140, 
1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 
1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140, 1140
), grade = c("3.5", "3.5", "3.75", "2.5", "3.75", "3.75", "3", 
"3.5", "4", "3.75", "3", "3.75", "3", "3", "3.25", "2.5", "3.25", 
"3.5", "2.25", "4", "2.5", "3.75", "0", "4", "3.75", "3.25", 
"4", "3.75", "2.75", "2.5", "3.75", "3.75", "3.5", "4", "3.25", 
"3.75", "3.5", "1.75", "4", "3", "2.75", "0", "3.75", "3.5", 
"3.75", "1.75", "3.5", "3.25", "3.25", "4")), class = c("tbl_df", 
"tbl", "data.frame"), row.names = c(NA, -50L))

grades_df
## a

grades_df |> group_by(section) |>
  summarise(mean_grade = mean(grade))

## b
## grade is <chr> not <dbl>
grades_df |> mutate(grade = as.double(grade)) |>
  group_by(section) |>
  summarise(mean_grade = mean(grade))

## c
ggplot(data = grades_df, aes(x = section, y = grade)) +
  geom_boxplot()

## d
grades_clean <- grades_df |>
  mutate(section = as.factor(section),
         grade = as.double(grade))

ggplot(data = grades_clean, aes(x = section, y = grade)) +
  geom_boxplot(outlier.shape = 8)
```

Class Exercise 4.

the name of the winning player: `<chr>` character
the surface that the match was played on (Hard, Grass, Clay): `<fct>`: factor.
the number of aces (unreturned serves) the winning player hit. `<int>`: integer.
the percentage of times the winning player made their first serve.: `<dbl>`: double.
whether or not the winning player was currently ranked higher than the losing player. `<lgl>` (or maybe factor).





