---
title: "Section 10: Joining"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

Class Exercise 0.

a. 5 rows.

b. 6 rows.

c. 8 rows.

d. 3 rows (A, C, E)

e. 3 rows.

f. 2 rows.

Class Exercise 1. To use bind_cols(), the two data frames must have cases in the same exact order.

Class Exercise 2.

```{r}
library(tidyverse)
library(here)
slumajors_df <- read_csv(here("data/SLU_Majors_17_21.csv"))
collegemajors_df <- read_csv(here("data/college-majors.csv"))
slumajors_df |> glimpse()
collegemajors_df |> glimpse()

left_join(slumajors_df, collegemajors_df,
          join_by(Major == Major))
## this is a problem: when joining, the keys have to match
## exactly.
```

Class Exercise 3.

```{r}
library(tidyverse)
library(here)
mortality_df <- read_csv(here("data/gun_violence_us.csv"))
mortality_df

dc_df <- tibble(state = "DC",
                mortality_rate = 16.7, ownership_rate = .087,
                region = "NE")
all_df <- bind_rows(mortality_df, dc_df)
all_df
```

Class Exercises 4-8

```{r}
df1 <- as_tibble(state.x77)
df2 <- as_tibble(state.abb)
df1
df2
states_df <- bind_cols(df1, df2) ## assuming the two data sets have the states in the same order

## ce 5
left_join(all_df, states_df, join_by(state == value))
## OR
full_join(all_df, states_df, join_by(state == value))

## ce 6
left_join(states_df, all_df, join_by(value == state))
## OR
inner_join(states_df, all_df, join_by(value == state))
```


```{r}
ne_df <- all_df |> filter(region == "NE")
semi_join(states_df, ne_df, join_by(value == state))

## ce 8
notne_df <- all_df |> filter(region != "NE")
anti_join(states_df, notne_df, join_by(value == state))
```




Your Turn 1.

a. true

b. true. For both a and b, an inner_join only returns rows with matching keys in both data frames while a left_ and full_ join return at least these rows and possibly others.

c. false: see YT 2 a and e for a counterexample.

Your Turn 2.

a. 6 rows (A, A, C, C, F, F)

b. 8 rows (A, A, B, C, C, D, D, D)

c. 10 rows (A, A, B, C, C, D, D, D, F, F)

d. 4 rows (A, A, C, C).

e. 4 rows (A, A, C, C, yvar is not kept at all)

f. 2 rows (F, F).