---
title: "Section 16: Intro to SQL"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(RSQLite)
library(DBI) 
library(dplyr)
library(dbplyr)

imdb_db <- dbConnect(RSQLite::SQLite(), 
                     dbname = here::here("data/imdb_tv.sqlite"))
imdb_db
```

```{r}
## lists the data tables (data frames) in our data base
dbListTables(conn = imdb_db)

## lists the fields (variables) in a particular data table
dbListFields(conn = imdb_db, name = "episode_rating")
```

```{r}
library(tidyverse)

sql <- "
  SELECT parentTconst, averageRating, numVotes
  FROM show_ratings 
  WHERE numVotes > 100000
"

## exercises 1 and 2
## WHERE is SQL's version of filter()
## FROM show_ratings says to do these operations on the
## show_ratings data table.
## SELECT is similar to select()

dbGetQuery(imdb_db, sql) |>
  as_tibble()
```

## Section 16.2: `dbplyr`

```{r}
library(dbplyr)
episode_rating_tbl <- tbl(imdb_db, "episode_rating")
episode_rating_tbl
## stores the episode_rating data table as its own object
```

```{r}
episode_query1 <- episode_rating_tbl |> 
  filter(averageRating > 9.0) |> 
  select(parentTconst, Episode_Name, numVotes, averageRating)
episode_query1
## data tables only show a "preview" of the output, so we don't
## know the total number of rows (??)
## the second line in the output clarifies where this database is.

episode_query1 |>
  collect()
## if we want to see all of the rows and we want to convert this
## from a data table to a data frame

episode_query1 |>
  show_query()
## shows the equivalent SQL commands for the dplyr query we made above.
```

```{r}
mean_query1 <- episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(mean_rating = mean(averageRating)) |>
  arrange(desc(mean_rating))
mean_query1

mean_query1 |> show_query()
```

Exercise 3.

```{r}
n_episode_query <- episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(n_episodes = n())

n_episode_query |> show_query()
## group_by()  is similar to GROUP BY
## summarise() is appearing is SELECT, where COUNT(*) is equivalent to n()
## and AS `n_episodes` is the name of the summary variable
```

## Section 16.3: SQL Matches to `dplyr`

```{r}
episode_rating_tbl |> show_query()
```

* `SELECT` in `SQL` covers `select()`, `rename()` (oldname AS newname), `mutate()` (x * y as prod), `summarise()` (MEDIAN(x) as medx).
* `GROUP BY` matches to `group_by()`.
* `WHERE` matches to `filter()`
* `ORDER BY` matches to `arrange()`

```{r}
show_info_tbl <- tbl(imdb_db, "show_info")

left_join(episode_rating_tbl, show_info_tbl,
          join_by(parentTconst == parentTconst)) |>
  show_query()
```

* LEFT JOIN is translated to `left_join()`: the key is specified as datatable1.key = datatable2.key instead of `join_by(key == key)`.

Exercise 4.

```{r}
show_info_tbl |> 
  filter(Show_Name == "Ozark") |>
  show_query()
```

Class Exercise 1.

```{r}
show_info_tbl |>
  rename(show_name = Show_Name) |>
  arrange(desc(startYear)) |>
  show_query()
```


Your Turns 1 through 4.

Your Turn 1.

```{r}
episode_rating_tbl |>
  group_by(parentTconst) |>
  summarise(best_episode = max(averageRating, na.rm = TRUE)) |>
  show_query()
## summary functions in SELECT like max, min, median, etc. get
## translated to summarise() while raw calcualtions like
## x + y, x / y, etc., get translated to mutate()
```

Your Turn 2.

```{r}
#| error: true
episode_rating_tbl |> mutate(test = fct_reorder(parentTconst,
                                                averageRating))
## errors: specific r packages like stringr, forcats, and lubridate
## do not translate into other languages
```

Your Turn 3.

```{r}
episode_rating_tbl |> filter(seasonNumber != "1") |>
  show_query()
## != is !=
episode_rating_tbl |>
  filter(seasonNumber %in% c("1", "2")) |>
  show_query()
## %in% is IN
episode_rating_tbl |> distinct(parentTconst) |>
  show_query()
## distinct() is DISTINCT
```


Your Turn 4.

```{r}
episode_rating_tbl |>
  filter(seasonNumber == "1") |>
  group_by(parentTconst) |>
  summarise(n_episodes = n()) |>
  arrange(desc(n_episodes)) |>
  show_query()
```

Your Turn 5.

```{r}
imdb_df <- left_join(episode_rating_tbl, show_info_tbl,
          join_by(parentTconst == parentTconst)) |>
  collect()
table(imdb_df$Show_Name)

imdb_stranger <- imdb_df |> filter(Show_Name == "Stranger Things") |>
  mutate(episodeNumber = as.numeric(episodeNumber))
ggplot(imdb_stranger, aes(x = episodeNumber,
                          y = averageRating)) +
  geom_line() +
  facet_wrap(~ seasonNumber )
```



