---
title: "Final Project"
format: revealjs
---

```{r}
#| echo: false
#| output: false
#| warning: false

library(schrute)
library(tidyverse)

schrute::theoffice

library(tidytext)
office_unnest <- theoffice |>
  unnest_tokens(output = "word", input = "text")
office_unnest

## stop_words gets loaded in when you load
## in the tidytext package
office_stop <- anti_join(office_unnest, stop_words,
                          by = join_by(word == word))

office_sum <- office_stop |> 
  filter(character == "Michael") |>
  group_by(word) |>
  summarise(n = n()) |>
  arrange(desc(n)) |>
  slice(1:20) |>
  mutate(word = fct_reorder(word, n))
```

## Project Goal

* What question(s) are you trying to answer with your project?

## Data Background

either use bullets to describe different variables OR

show the first couple of rows and a few columns of interest

## Table or Graph of Results

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}

ggplot(data = office_sum, aes(x = word, y = n)) +
  geom_segment(aes(xend = word, y = 0, yend = n)) +
  geom_point() +
  theme_minimal() +
  coord_flip()
```

## Table or Graph of Results
